\textbf{\sc Deliberative Autonomy for Marine Robotics}

### 2012-01-25 trunk/arch.tex
added a little paragraph presenting T-REX as conceptualization of every reactor as deliberation processes

@@

>>>>

{\em\color{red} Following is an attempt to present TREX as a big
  deliberation machine}
The overall execution cycle of the agent is also abstracted as some
continuous planning/deliberation cycle between all the
reactors. Indeed , at the agent interface level, each reactor provided
only two abstract calls that conceptualize their execution:
\begin{itemize}
\item \texttt{synchronize} that takes the last {\em external}
  observations as argument and will return the {\em internal}
  observations for this reactor or a failure report.
\item \texttt{step} that accept new goals requested (if any) on the
  {\em internal} timeline of this reactor and will execute one step of
  deliberation. The returned value indicates if a complete plan has
  been found and in that case will also provide the part of this plan
  applied to {\em external} timelines of this reactor.
\end{itemize}
Both of these two deliberative functions have different scope and
temporal constraints. The \texttt{synchronize} call is considered as
atomic and will focus only on state inference from the current tick by
identifying the current value of the {\em internal} state variables
from the latest updates on the {\em external} variables. This process
will then propagate up in the reactor graph at any single tick to
ensure that all reactors are aware of the current state of the
world. \texttt{step} relates to a single decision step of the planning
process which is focused on the future evolution of the state
variables and it can take multiple steps spanning other multiple
ticks for a reactor to produce a plan that is both complete and
valid. when such plan is identified its {\em external} part is then
dispatched to the corresponding reactors which in turn can start their
deliberation \texttt{step}s in order to satisfy these new
objectives. as a consequence physical execution will occur as these
goals and planning propagate down in the reactor graph until they
eventually become goals for the lowest level reactor which -- instead
of deliberating -- may just transform this goal in a physical command
execution. 
{\em\color{red} end of the attempt; need to rework the paragraphs
  below accordingly}

@@

The planning process inside a reactor is directly based on the way
europa work which was described in previous section. Still some simple
alterations were made in  order to both include deliberation
information one reactor provide to the architecture -- such as its
deliberation latency and its planning look-ahead both expressed in 
ticks -- and the fact that we potentially need to interleave this
deliberation with not the synchronization process but also with 
the deliberation of other reactors\footnote{Current implementation of
  T-REX is running on a single process and it is the responsibility of
  the agent itself to emulate the multi-threading of the reactors
  deliberation and synchronization}. For this reason we 
decompose the algorithm \ref{alg:europa:solve} inside the reactor by
having the recursive calls decomposed in single steps. This step
function will be called after synchronization and as long as the
previous call in this tick did not return \textsf{true} -- which would
express that the current solution is complete.
{\em will need to change that in algorithm environment and refine it}

\begin{verbatim}
bool step(PartialPlan plan, DecisionStack stack) 
   if( isInconsistent(plan) ) {
     relaxPlan(plan, stack);
     notify("No plan");
     return true;
   } 
   Flaw flaw = chooseFlaw(flaw);
   if( flaw==NULL )
     return true;
   DecisionPoint decision = makeDecisionPoint(flaw, plan);
   while(1) {
     while( decision.hasNext() ) {
       PartialPlan pp = decision.executeNext();
       if( isInconsistent(pp) ) {
         decision.undo();
       }  else {
          plan = pp;
          stack.push(decision);
          return chooseFlaw(flaw)==NULL;
       }
     }
     if( decisionStack.empty() ) {
       notify("No plan");
       return true;
     } else {
        repeat {
          decision = decisionStack.pop();
          decision.undo();
        } until( decision.hasNext() );
        return false;
     }
   } 
}
\end{verbatim}

One critical aspect of this deliberation step is to ensure that at the
exit of this call the $plan$ is not proven inconsistent. Indeed, it is
always possible that the next call made by the agent is the
synchronization which would immediately fail if the $plan$ is
inconsistent before any solving attempt made for synchronization. 
For this reason if the plan is proven inconsistent during a
deliberation step we either backtrack in the decision stack until we
find a node fo the tree that have remaining alternate solution or
simply fully relax the plan in situation there's absolutely no other
alternative  (by removing all the goals and keeping
only recent observations produced). These constraints are yet again
related to the fact that for the agent the only crucial functionality
of a reactor is to resolve successfully its synchronization.

The number of flaw between two steps can evolve for the following
reasons that are exogenous to deliberation :
\begin{enumerate}
\item A new goal has been posted on one of the reactor timeline. This
  is usually due to another reactor finding a plan and posting the
  corresponding new objective to this reactor.
\item The outcome of synchronization created new flaws that did not
  need to be resolved during synchronization.
\item A previously requested goal has be recalled by the initial
  requester. This can happen when the reactor did identify that is
  initial plan is no longer valid. For example as synchronization
  proved this plan to be inconsistent.
\end{enumerate}
These 3 events will potentially produce new flaws to be resolved while
any single deliberation step will usually reduce the number of flaws
to be resolved.

The scope of planning for the reactor plays also a role in the
deliberation process as we limit deliberation to the sliding temporal
window $[\tau+\lambda, \tau+\lambda+\pi]$ where:
\begin{itemize}
\item $\tau$ is the current tick
\item $\lambda$ is the specified latency of this reactor. Which is an
  indicator on the expected maximum time this reactor is expected
  before being able to produce a plan.\footnote{Do note that this
    parameter remains indicative and, while it is highly recommended to
    select a sound value, a failure to produce a plan in this delay is
    not considered as a critical failure of the reactor}
\item $\pi$ is the planning look-ahead of the reactor. which indicates
  how many ticks in the future this reactor is looking ahead while planning.
\end{itemize}
This scope allow to filter out any tokens that are either necessarily
ending before $\tau$ or necessarily starting after. Focusing the
[planning problem to only the tokens that are in a reasonable
future. This window is taken into account by the agent which will
notify the reactor of new {\em internal} goals only when they overlap
this window but also at the deliberation solver in order to reduce the
number of tokens to be evaluated and consequently reduce the cost of
deliberation. As time advance flaws that where initially beyond this
window will become active flaw that the reactor can then evaluate
resulting on an apparent continuous planning of the reactor.

When there's no more flaw is present in the current planning scope a
plan  solution is found. This has two effects:
\begin{enumerate}
\item This reactor does not need to deliberate until the next synchronization.
\item The {\em external} part of this plan can then be posted which
  eventually (at the end of the current tick if these goals overlap
  the owner planning window) will be dispatched as goals triggering
  new deliberation on these reactors.
\end{enumerate}
